Taking Dyads Seriously

### 2020-06-16 paper/3_sim.tex
wrapping up

Through a simulation study, we highlight the utility of AME as an inferential tool for dyadic analysis.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM). While both these models have led to numerous contributions to a variety of literatures, the applicability of these approaches may be limited to certain types of networks and individual level characteristics. Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable. The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.\footnote{\textcolor{blue}{The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria. With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l). In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).}} Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}

@@

The simulation shows that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is to directly estimate unobserved dependencies through the random effects structure of the model. Scholars can use this framework in an iterative fashion: beginning with an estimated model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their model.  \textcolor{blue}{Importantly, this simulation underscores how a careful consideration of a systems' interconnectedness, both through theoretical approaches and empirical models, can result in more precise estimates of direct and indirect effects across the system.}

>>>>

Through a simulation study, we highlight the utility of AME as an inferential tool for dyadic analysis.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM). While both these models have led to numerous contributions to a variety of literatures, the applicability of these approaches may be limited to certain types of networks and individual level characteristics. Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable. The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.\footnote{The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria. With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l). In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).} Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}

@@

The simulation shows that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is to directly estimate unobserved dependencies through the random effects structure of the model. Scholars can use this framework in an iterative fashion: beginning with an estimated model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their model. Importantly, this simulation underscores how a careful consideration of a systems' interconnectedness, both through theoretical approaches and empirical models, can result in more precise estimates of direct and indirect effects across the system.
